Home Nighttime Cough Monitoring

### 2024-10-17 Lilah Melzer/cough_monitor_analysis.qmd
mixed model with random slope for Lilah

```{r data cleaning}
#| message: false
#| warning: false
# Import REDCap data
source(paste0(github_dir, "/CF/Lilah Melzer/cough_monitor_import.r"))
# Import use data
use <- read.csv("./Data_Clean/Cough study vacation days consent date.csv",
  na.strings = ""
)
# Import counts
agg_counts <- read.csv(
  "./Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
  na.strings = c("", "NMD"), header = F
)
# Format aggregated count data
agg_counts <- data.frame(t(agg_counts))
colnames(agg_counts) <- agg_counts[1, ]
agg_counts$`Date/Device` <- NULL
agg_counts <- agg_counts[-1, ]
colnames(agg_counts)[1] <- "sid"
agg_counts$sid <- sub("FN", "", agg_counts$sid)
# Wide to long, fix dates
agg_counts <- agg_counts %>%
  pivot_longer(
    cols = -one_of("sid"), names_to = "Date", values_to = "Coughs"
  ) %>%
  mutate(
    Date = mdy(Date),
    Coughs = as.numeric(Coughs)
  )
# Add start date
agg_counts$sid <- as.numeric(agg_counts$sid)
agg_counts$Start <-
  mdy(use$Start.date[match(agg_counts$sid, use$Participant.number)])
# Count number of nights
agg_counts <- agg_counts %>%
  filter(Date >= Start) %>%
  group_by(sid) %>%
  mutate(Night = row_number()) %>%
  ungroup()
# Demographics data
data$Diagnosis <- factor(sub(".*arm_*", "", data$redcap_event_name),
  levels = 1:2,
  labels = c("CF", "Non-CF")
)
demo <- data %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  select(
    sid, Diagnosis, age_at_consent, height, weight, pre_fev,
    nih_sex.factor:nih_ethnicity.factor, mutation1.factor, mutation2.factor,
    cftr_mod.factor, cftr_mod_type.factor, mucolytic.factor:antih.factor
  )
# Calculate FEV1% - rspiro not accepting NA values for gender, which is really
# annoying
fev1 <- demo %>%
  select(age_at_consent, height, nih_sex.factor, pre_fev) %>%
  filter(nih_sex.factor %in% c("Male", "Female"))
fev1$pre_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
  factor(fev1$nih_sex.factor, levels = c("Male", "Female")),
  FEV1 = fev1$pre_fev
)
demo <- left_join(demo, fev1[, c("sid", "pre_fev_pp")])
label(demo$pre_fev_pp) <- "Pre-bronchodilator FEV1 % Predicted"
# CF genotype
demo$Genotype <-
  factor(rowSums(demo[, c("mutation1.factor", "mutation2.factor")] == "F508del"),
    levels = c(2, 1, 0),
    labels = c("F508del homozygous", "F508del heterozygous", "No F508del")
  )
demo[, grep("mutation", colnames(demo))] <- NULL
# Add group to counts
agg_counts$Diagnosis <-
  demo$Diagnosis[match(agg_counts$sid, as.numeric(as.character(demo$sid)))]
# Per night
per_night <- agg_counts %>%
  group_by(sid) %>%
  summarise(
    median_coughs = median(Coughs, na.rm = T),
    nights_data = sum(!is.na(Coughs))
  )
demo$sid <- as.numeric(demo$sid)
demo <- left_join(demo, per_night)
label(demo$median_coughs) <- "Median Coughs per Night"
label(demo$nights_data) <- "Nights With Cough Data"
# Set table options for the document
mycontrols <-
  tableby.control(numeric.stats = c("Nmiss", "medianq1q3", "range"),
                  numeric.test = "kwt")
```

@@

```{r}
mod <- glmer(Coughs ~ Night * Diagnosis + (1 | sid),
  data = agg_counts, family = "poisson"
)
gtsummary::tbl_regression(mod, exponentiate = T)

mod <- glmer(Coughs ~ Diagnosis + (1 | sid),
  data = agg_counts, family = "poisson"
)
gtsummary::tbl_regression(mod, exponentiate = T)
```

>>>>

```{r data cleaning}
#| message: false
#| warning: false
# Import REDCap data
source(paste0(github_dir, "/CF/Lilah Melzer/cough_monitor_import.r"))
# Import use data
use <- read.csv("./Data_Clean/Cough study vacation days consent date.csv",
  na.strings = ""
)
# Import counts
agg_counts <- read.csv(
  "./Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
  na.strings = c("", "NMD"), header = F
)
# Format aggregated count data
agg_counts <- data.frame(t(agg_counts))
colnames(agg_counts) <- agg_counts[1, ]
agg_counts$`Date/Device` <- NULL
agg_counts <- agg_counts[-1, ]
colnames(agg_counts)[1] <- "sid"
agg_counts$sid <- sub("FN", "", agg_counts$sid)
# Wide to long, fix dates
agg_counts <- agg_counts %>%
  pivot_longer(
    cols = -one_of("sid"), names_to = "Date", values_to = "Coughs"
  ) %>%
  mutate(
    Date = mdy(Date),
    Coughs = as.numeric(Coughs)
  )
# Add start date
agg_counts$sid <- as.numeric(agg_counts$sid)
agg_counts$Start <-
  mdy(use$Start.date[match(agg_counts$sid, use$Participant.number)])
# Count number of nights
agg_counts <- agg_counts %>%
  filter(Date >= Start) %>%
  group_by(sid) %>%
  mutate(Night = row_number()) %>%
  ungroup()
# Demographics data
data$Diagnosis <- factor(sub(".*arm_*", "", data$redcap_event_name),
  levels = 1:2,
  labels = c("CF", "Non-CF")
)
demo <- data %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  select(
    sid, Diagnosis, age_at_consent, height, weight, pre_fev,
    nih_sex.factor:nih_ethnicity.factor, mutation1.factor, mutation2.factor,
    cftr_mod.factor, cftr_mod_type.factor, mucolytic.factor:antih.factor
  )
# Calculate FEV1% - rspiro not accepting NA values for gender, which is really
# annoying
fev1 <- demo %>%
  select(age_at_consent, height, nih_sex.factor, pre_fev) %>%
  filter(nih_sex.factor %in% c("Male", "Female"))
fev1$pre_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
  factor(fev1$nih_sex.factor, levels = c("Male", "Female")),
  FEV1 = fev1$pre_fev
)
demo <- left_join(demo, fev1[, c("sid", "pre_fev_pp")])
label(demo$pre_fev_pp) <- "Pre-bronchodilator FEV1 % Predicted"
# CF genotype
demo$Genotype <-
  factor(rowSums(demo[, c("mutation1.factor", "mutation2.factor")] == "F508del"),
    levels = c(2, 1, 0),
    labels = c("F508del homozygous", "F508del heterozygous", "No F508del")
  )
demo[, grep("mutation", colnames(demo))] <- NULL
# Add group to counts
agg_counts$Diagnosis <-
  demo$Diagnosis[match(agg_counts$sid, as.numeric(as.character(demo$sid)))]
# Per night
per_night <- agg_counts %>%
  group_by(sid) %>%
  summarise(
    median_coughs = median(Coughs, na.rm = T),
    nights_data = sum(!is.na(Coughs))
  )
demo$sid <- as.numeric(demo$sid)
demo <- left_join(demo, per_night)
label(demo$median_coughs) <- "Median Coughs per Night"
label(demo$nights_data) <- "Nights With Cough Data"
# Set table options for the document
mycontrols <-
  tableby.control(
    numeric.stats = c("Nmiss", "medianq1q3", "range"),
    numeric.test = "kwt"
  )
```

@@

```{r}
mod <- glmer(Coughs ~ Night + Diagnosis + (Night | sid),
  data = agg_counts, family = "poisson"
)
gtsummary::tbl_regression(mod, exponentiate = T)
```
